---
title: "Business Intelligence Lab Submission Markdown"
author: "naive"
date: "4/10/2023"
output:
  github_document: 
    toc: yes
    toc_depth: 4
    fig_width: 6
    fig_height: 4
    df_print: default
editor_options:
  chunk_output_type: console
---

# Student Details

+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Student ID Numbers and Names of Group Members** | *\<list one student name, class group (just the letter; A, B, or C), and ID per line, e.g., 123456 - A - John Leposo; you should be between 2 and 5 members per group\>* |
|                                                   |                                                                                                                                                                          |
|                                                   | 1.  135575 - B - Dennis Nzioki.                                                                                                                                             |
|                                                   |                                                                                                                                                                          |
|                                                   | 2.  134645 - B - Vivean Lydiah                                                                                                                                             |
|                                                   |                                                                                                                                                                          |
|                                                   | 3.  134765 - B - Nicholas Munene                                                                                                                                             |
|                                                   |                                                                                                                                                                          |
|                                                   | 4.  131653- B - Terry Joan                                                                                                                                             |
|                                                   |                                                                                                                                                                          |
|                                                   | 5.  124428 - B - Eston Gichuhi                                                                                                                                              |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **GitHub Classroom Group Name**                   | *\<specify the name of the team you created on GitHub classroom\>*                                                                                                       |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Course Code**                                   | BBT4206                                                                                                                                                                  |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Course Name**                                   | Business Intelligence II                                                                                                                                                 |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Program**                                       | Bachelor of Business Information Technology                                                                                                                              |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| **Semester Duration**                             | 21^st^ August 2023 to 28^th^ November 2023                                                                                                                               |
+---------------------------------------------------+--------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

# Setup Chunk

We start by installing all the required packages
We start by installing all the required packages

```{r Install Packages, echo=TRUE, message=FALSE, warning=FALSE}
## formatR - Required to format R code in the markdown ----
if (!is.element("formatR", installed.packages()[, 1])) {
  install.packages("formatR", dependencies = TRUE,
                   repos="https://cloud.r-project.org")
}
require("formatR")
if (require("caret")) {
  require("caret")
} else {
  install.packages("caret", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## klaR ----
if (require("klaR")) {
  require("klaR")
} else {
  install.packages("klaR", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## e1071 ----
if (require("e1071")) {
  require("e1071")
} else {
  install.packages("e1071", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## readr ----
if (require("readr")) {
  require("readr")
} else {
  install.packages("readr", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## LiblineaR ----
if (require("LiblineaR")) {
  require("LiblineaR")
} else {
  install.packages("LiblineaR", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
}

## naivebayes ----
if (require("naivebayes")) {
  require("naivebayes")
} else {
  install.packages("naivebayes", dependencies = TRUE,
                   repos = "https://cloud.r-project.org")
     install.packages("mlbench", dependencies = TRUE)  #nolint
  
}
if (!is.element("mlbench", installed.packages()[, 1])) {}
  require("mlbench") #nolint
```

------------------------------------------------------------------------

**Note:** the following "*KnitR*" options have been set as the defaults in this markdown:\
`knitr::opts_chunk$set(echo = TRUE, warning = FALSE, eval = TRUE, collapse = FALSE, tidy.opts = list(width.cutoff = 80), tidy = TRUE)`.

More KnitR options are documented here <https://bookdown.org/yihui/rmarkdown-cookbook/chunk-options.html> and here <https://yihui.org/knitr/options/>.

```{r setup, echo=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
	eval = TRUE,
	echo = TRUE,
	warning = FALSE,
	collapse = FALSE,
	tidy = TRUE
)
```
# DATASET 1 (Splitting the dataset): PimaIndians Diabetes
```{r load dataset}
data("PimaIndiansDiabetes")
summary(PimaIndiansDiabetes)
str(PimaIndiansDiabetes)
```
##Split the dataset:
```{r split dataset}
train_index <- createDataPartition(PimaIndiansDiabetes$diabetes,
                                   p = 0.75,
                                   list = FALSE)
PimaIndiansDiabetes_train <- PimaIndiansDiabetes[train_index, ]
PimaIndiansDiabetes_test <- PimaIndiansDiabetes[-train_index, ]
```
##Train a Naive Bayes classifier using the training dataset
this can be done by using an naive bayes function in the e1071 package or the caret package
```{r train naive}
PimaIndiansDiabetes_model_nb_e1071 <- # nolint
  e1071::naiveBayes(diabetes ~ pregnant + glucose + pressure + triceps + insulin + mass +
                      pedigree + age,
                    data = PimaIndiansDiabetes_train)
PimaIndiansDiabetes_model_nb_caret <- # nolint
  caret::train(diabetes ~ ., data =
                 PimaIndiansDiabetes_train[, c("pregnant", "glucose", "pressure", "triceps", "insulin",
                                               "mass", "pedigree", "age", "diabetes")],
               method = "naive_bayes")
```
##Test the trained model using the testing dataset
### 3.a. Test the trained e1071 Naive Bayes model using the testing dataset ----
```{r}
predictions_nb_e1071 <-
  predict(PimaIndiansDiabetes_model_nb_e1071,
          PimaIndiansDiabetes_test[, c("pregnant", "glucose", "pressure", "triceps", "insulin",
                                               "mass", "pedigree", "age", "diabetes")])
```
### 3.b. Test the trained caret Naive Bayes model using the testing dataset ----
```{r}
predictions_nb_caret <-
  predict(PimaIndiansDiabetes_model_nb_caret,
          PimaIndiansDiabetes_test[, c("pregnant", "glucose", "pressure", "triceps", "insulin",
                                       "mass", "pedigree", "age", "diabetes")])
```
## 4. View the Results ----
### 4.a. e1071 Naive Bayes model and test results using a confusion matrix ----

```{r}
print(predictions_nb_e1071)
caret::confusionMatrix(predictions_nb_e1071,
                       PimaIndiansDiabetes_test[, c("pregnant", "glucose", "pressure",
                                                  "triceps", "insulin", "mass",
                                                  "pedigree",
                                                  "age",
                                                  "diabetes" 
                                                  )]$diabetes)
plot(table(predictions_nb_e1071,
           PimaIndiansDiabetes_test[, c("pregnant", "glucose", "pressure",
                                        "triceps", "insulin", "mass",
                                        "pedigree",
                                        "age",
                                        "diabetes" 
           )]$diabetes))
```
### 4.b. caret Naive Bayes model and test results using a confusion matrix ----
```{r}
print(PimaIndiansDiabetes_model_nb_caret)
caret::confusionMatrix(predictions_nb_caret,
                       PimaIndiansDiabetes_test[, c("pregnant", "glucose", "pressure",
                                                    "triceps", "insulin", "mass",
                                                    "pedigree",
                                                    "age",
                                                    "diabetes" 
                       )]$diabetes)
plot(table(predictions_nb_caret,
           PimaIndiansDiabetes_test[, c("pregnant", "glucose", "pressure",
                                        "triceps", "insulin", "mass",
                                        "pedigree",
                                        "age",
                                        "diabetes" 
           )]$diabetes))
```
## 5. Classification: SVM with Repeated k-fold Cross Validation ----

### 5.a. SVM Classifier using 5-fold cross validation with 3 reps ----
the support vector machine is trained with the purpose of classification using 'diabetes variable. we will use a 5-fold cross-validation train control with 3 reps
```{r}
train_control <- trainControl(method = "repeatedcv", number = 5, repeats = 3)

PimaIndiansDiabetes_model_svm <-
  caret::train(`diabetes` ~ ., data = PimaIndiansDiabetes_train,
               trControl = train_control, na.action = na.omit,
               method = "svmLinearWeights2", metric = "Accuracy")
```
### 5.b. Test the trained SVM model using the testing dataset ----
```{r}
predictions_svm <- predict(PimaIndiansDiabetes_model_svm, PimaIndiansDiabetes_test[, 1:9])
```
### 5.c. View a summary of the model and view the confusion matrix ----
```{r}
print(PimaIndiansDiabetes_model_svm)
caret::confusionMatrix(predictions_svm, PimaIndiansDiabetes_test$diabetes)
```
## 6. Classification: Naive Bayes with Leave One Out Cross Validation----
In Leave One Out Cross-Validation (LOOCV), a data instance is left out and a model constructed on all other data instances in the training set. This is repeated for all data instances.
### 6.a. Train a Naive Bayes classifier based on an LOOCV ----
```{r}
train_control <- trainControl(method = "LOOCV")

PimaIndiansDiabetes_model_nb_loocv <-
  caret::train(`diabetes` ~ ., data = PimaIndiansDiabetes_train,
               trControl = train_control, na.action = na.omit,
               method = "naive_bayes", metric = "Accuracy")
```
### 6.b. Test the trained model using the testing dataset ====
```{r}
predictions_nb_loocv <-
  predict(PimaIndiansDiabetes_model_nb_loocv, PimaIndiansDiabetes_test[, 1:9])
```
### 6.c. View the confusion matrix ====
```{r}
print(PimaIndiansDiabetes_model_nb_loocv)
caret::confusionMatrix(predictions_nb_loocv, PimaIndiansDiabetes_test$diabetes)
```



